The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}

### 2018-08-28 submission/GRB111117A.tex
Published

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
%\documentclass[referee]{aa}

@@

\usepackage{url}
\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0.3,0.7}
\hypersetup{colorlinks=true,
	linkcolor=linkcolor, 
	citecolor=linkcolor,
	filecolor=linkcolor, 
	urlcolor=linkcolor}

@@

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy and the physical parameters of the GRB. The rest-frame
X-ray derived hydrogen column density, for example, is the highest compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs.

@@

There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs) come
from the merger of a neutron star (NS), either with another NS, or a black hole,
due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. This association was
recently confirmed by the simultaneous and co-spatial detection of gravitational
waves (GWs) from a binary neutron star merger and a sGRB
\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}. The
absence of associated supernovae in deep searches
\citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011} supports this idea and
distinguishes the physical origin of sGRBs from their long-duration
counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

@@

We here present a spectrum of the host galaxy of the short GRB~111117A
($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is
significantly higher than the previously estimated redshift based on photometric
studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest frame
properties based on this new distance compared to previous analyses and revisit
the host properties derived from the new solution to the spectral energy
distribution (SED) fit. While no optical afterglow was detected, the excellent
localization from a detection of the X-ray afterglow by the \emph{Chandra} X-ray
Observatory allows us to discuss the positioning and environmental properties of
this remarkably distant sGRB. We use the $\Lambda$CDM cosmology parameters
provided by \citet{Planck2015} in which the universe is flat with $H_0 =
67.7$\,km\,s$^{-1}$ Mpc$^{-1}$ and $\Omega_\mathrm{m} = 0.307$. All magnitudes
are given in the AB system.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
	\caption{
	FORS2 $R$-band imaging of the field of GRB~111117A with the X-shooter slit
overlaid. The slit position represents 4 epochs of spectroscopic observations
taken at similar position angles. The corresponding photometry is shown in
Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as derived
from the \emph{Chandra} observations in \citet{Sakamoto2013}.
	}
	\label{fig:spec_setup}
\end{figure}

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
%	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
%	\label{fig:line}
%\end{figure}

\begin{figure*}
	\centering
	\includegraphics[width=18cm]{figures/grid.pdf}
	\caption{2D-images of the emission lines corresponding to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha. The location of bright skylines are marked by blue boxes. The locations of the emission lines are indicated with red ellipses. Because the host is observed in nodding-mode, negative images of the emission lines appear on both sides in the spatial direction. For the upper, left panel containing \lya, the systemic redshift position of \lya{} is marked by a yellow, vertical, dashed line. The red part of the \oii$\lambda$3727-doublet is affected by atmospheric absorption.}
	\label{fig:line}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
	\caption{Best-fit SED to the derived photometry. The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes. Slit losses has been corrected for based on the average seeing of the observations, as confirmed by the comparison with the photometry. The blue, dashed line is the corresponding error spectrum, smoothed for presentation purposes. The reason for the spectral gaps at 5500 \AA{} and 10000 \AA{} is from the merging of the arms.}
	\label{fig:SED}
\end{figure*}

@@

\input{tables/spec_overview.tex}

@@

\begin{figure}
	\centering \includegraphics[width=9cm]{figures/photometry_mosaic.pdf}
\caption{Mosaic showing all used imaging. The host of GRB~111117A is marked by
	a red circle with a 2\farc0 radius. This is the same size as the aperture used
	for deriving the photometry. Each panel is 20\arcsec~in size, north is up, east
	is left. Worth noting is the relative depth of the GMOS-N and FORS2 $z$-band
	images. For reference, the object located to the south-east of the host and
	marked with a blue, dashed circle has an extinction corrected magnitude of
	$23.11 \pm 0.09$ ($23.10 \pm 0.18$) in the FORS2 (GMOS-N) $z$-band.}
\label{fig:photplot}
\end{figure}

\input{tables/phot_overview.tex}

@@

This securely makes GRB~111117A, by far, the highest redshift sGRB detected to
date. The redshift and classification of GRB~111117A imply that it occurred when
the universe was younger by $\sim$ 3 Gyr compared to any other securely
classified sGRB ever detected. If the merger of NSs is the primary agent for the
$r$-process element enrichment of the universe \citep{Goriely2011, Ji2016,
	Komiya2016}, this marks the earliest detection of this process.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
	\caption{Rest-frame XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shaded region is a compilation of long GRB lightcurves \citep{Evans2007, Evans2009} where the color represents density. The light blue lines are sGRB lightcurves from bursts with duration of $T_{90} \lesssim 2$ s and those that were classified as short in \citet{Kann2011, Berger2014, DAvanzo2014a}. The thick green line is GRB~111117A. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population. The insert shows the X-ray luminosity distributions of sGRBs and lGRBs at 10 ks, indicated by the vertical dashed line in the main panel.}
	\label{fig:sxray_lightcurve}
\end{figure}

@@

\begin{figure}
	\centering \includegraphics[width=9cm]{figures/NH_z.pdf} \caption{Rest frame,
	X-ray derived equivalent hydrogen column density of GRB~111117A compared to
	complete samples of both long and short populations of GRBs. The sample of
	lGRBs from \citet{Arcodia2016} is shown in red, where detections are also shown
	with a kernel density estimate of the points and the limits on $N_\mathrm{H,X}$
	are shown with arrows. The complete sample of sGRB by \citet{DAvanzo2014a} is
	shown in blue, where again the limits are indicated by arrows. Marginalizations
	over both axes are shown on the right and on the top of the plot, where the
	limits are shown as semi-transparent bars and detections as solid ones. The red
	curves in the marginalization plots are again the kernel density estimates of
	the \citet{Arcodia2016} sample.} \label{fig:NH_z}
\end{figure}

@@

A critical test to assess if the power law delay time models can be accommodated
by the current observation is to check if the implied sGRB rate at higher
redshift does not exceed the number of observed sGRBs without redshifts. Of the
100  sGRBs observed by \textit{Swift}, 20 have secure redshifts, and another 7
have a tentative redshift measurement\footnote{This is based on
	\url{http://www.astro.caltech.edu/grbox/grbox.php}, selecting all short,
	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning that $> 73$ per cent of all
sGRBs observed with \textit{Swift} are missing redshifts. More recently,
\citet{Fong2017} \citep{LIGOScientificCollaboration2017b} compiled a list of 36
(33) sGRBs with redshift measurements and, using this number, the redshift
incompleteness of sGRBs decreases to 64 per cent. Additional to the potential
number of high-$z$ event already detected but missing redshifts,
\citet{Behroozi2014} parametrized the \textit{Swift} redshift sensitivity and
found that the mean detection probability for sGRBs at $z \sim2$ was only $\sim1$
per cent of the mean detection probability at $z \sim1$, assuming that the
unknown beaming angle of sGRBs stays constant with time. What this means is that
at the present, there is almost no limit on the number of sGRBs that could be at
higher redshifts.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/prog_sep.pdf}
	\caption{Constraints on the initial progenitor separation, given binary constituent masses. As can be seen, a heavier system will inspiral faster, leading to a weaker constraint on the initial separation of the binary. The most probable value, the 68 per cent, and 98 percent posterior predictive intervals of the NS binary mass distribution from \citet{Kiziltan2013} are shown with the dashed ellipses, which corresponds to a constraint on maximal initial separation of $a_0 <3.1^{+0.2(+0.4)}_{-0.2(-0.4)}~R_\odot$, respectively.}
	\label{fig:prog_sep}
\end{figure}

@@

The delay time between formation and explosion is well accommodated by the
models of \citet{Belczynski2006}, although the longest delay times are excluded.
This is especially true given the late type nature of the host
\citep{OShaughnessy2008}. The same holds, if NS binaries are primarily formed
through dynamical interactions in globular clusters \citep{Lee2010, Church2011}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
\documentclass[referee]{aa}

@@

\usepackage{url}
\usepackage{color,hyperref}
%\definecolor{linkcolor}{rgb}{0,0.3,0.7}
%\hypersetup{colorlinks=true,
%	linkcolor=linkcolor, 
%	citecolor=linkcolor,
%	filecolor=linkcolor, 
%	urlcolor=linkcolor}

@@

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy and the physical parameters of the GRB. The rest-frame
X-ray derived hydrogen column density, for example, is the highest compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs. From the detection of \lya~emission in the
spectrum, we are able to constrain the escape fraction of \lya~in the host.

@@

There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs) come
from the merger of a neutron star (NS), either with another NS, or a black hole,
due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. This association was
recently confirmed by the simultaneous and co-spatial detection of gravitational
waves (GWs) from a binary neutron star merger and a sGRB
\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}, however,
to what degree the sGRB associated with GW170817 is a cosmological sGRB remains
an open question \citep{Lyman2018, Margutti2018}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b,
	Kann2011} supports this idea and distinguishes the physical origin of sGRBs from
their long-duration counterparts, \citep[albeit see also][]{Fynbo2006b,
	Valle2006, Gal-Yam2006}.

@@

We here present a spectrum of the host galaxy of the short GRB~111117A
($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is
significantly higher than the previously estimated redshift of $z \sim 1.3$ based on photometric
studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest frame
properties based on this new distance compared to previous analyses and revisit
the host properties derived from the new solution to the spectral energy
distribution (SED) fit. While no optical afterglow was detected, the excellent
localization from a detection of the X-ray afterglow by the \emph{Chandra} X-ray
Observatory allows us to discuss the positioning and environmental properties of
this remarkably distant sGRB. We use the $\Lambda$CDM cosmology parameters
provided by \citet{Planck2015} in which the universe is flat with $H_0 =
67.7$\,km\,s$^{-1}$ Mpc$^{-1}$ and $\Omega_\mathrm{m} = 0.307$. All magnitudes
are given in the AB system.

@@

\begin{figure}
	\centering
%	\includegraphics[width=9cm]{GRB111117A_spec_obs.pdf}
	\caption{
	FORS2 $R$-band imaging of the field of GRB~111117A with the X-shooter slit
overlaid. The slit position represents 4 epochs of spectroscopic observations
taken at similar position angles. The corresponding photometry is shown in
Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position (R.A., Dec.) (J2000) = (00$^{\rm h}$ 50$^{\rm m}$ 46.264$^{\rm s}$, +23$^{\circ}$ 00$^{\prime}$ 39.98$^{\prime\prime}$) as derived
from the \emph{Chandra} observations in \citet{Sakamoto2013}.
	}
	\label{fig:spec_setup}
\end{figure}

%\begin{figure}
%	\centering
%%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
%	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
%	\label{fig:line}
%\end{figure}

\begin{figure*}
	\centering
%	\includegraphics[width=18cm]{grid.pdf}
	\caption{2D-images of the emission lines corresponding to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha. The location of bright skylines are marked by blue boxes. The locations of the emission lines are indicated with red ellipses. Because the host is observed in nodding-mode, negative images of the emission lines appear on both sides in the spatial direction. For the upper, left panel containing \lya, the systemic redshift position of \lya{} is marked by a yellow, vertical, dashed line. The red part of the \oii$\lambda$3727-doublet is affected by atmospheric absorption.}
	\label{fig:line}
\end{figure*}

\begin{figure*}
	\centering
%	\includegraphics[width=16cm]{SEDspecphot.pdf}
	\caption{Best-fit SED to the derived photometry. The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes. Slit losses has been corrected for based on the average seeing of the observations, as confirmed by the comparison with the photometry. The blue, dashed line is the corresponding error spectrum, smoothed for presentation purposes. The reason for the spectral gaps at 5500 \AA{} and 10000 \AA{} is from the merging of the arms.}
	\label{fig:SED}
\end{figure*}

@@

\begin{table*}[!ht]

	\centering
	\caption{Overview of the spectroscopic observations. ``JH'' in the slit width refers to observations where a $K$-band blocking filter has been used. The seeing is determined from the width of the spectral trace of a telluric standard star, observed close in time to the host observation. The spectral resolution, $R$, is measured from unresolved telluric absorption lines in the spectrum of the telluric standard star. \label{tab:spec_overview}}
	\begin{tabular}{cccccccc}
		\hline\hline
		{Observation epoch} &  \multicolumn{3}{c}{Exposure time (s)} & Slit width & Airmass & Seeing & $R$   \\ [1.5pt]
		\hline
		(UT) & UVB  & VIS & NIR &  (arcsec)   & {} & (arcsec)  & {VIS/NIR}  \\ [1.5pt]
		\hline
		2011-11-19 01:33 & 2 $\times$ 2400 & 2 $\times$ 2400 & 8 $\times$ 600 & 1.0/1.0/0.9 & 1.49 & 0.75 & 11600/6700 \\
		2013-07-15 09:02 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.53 & 0.98 & 9600/8900 \\
		2013-08-03 07:37 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.55 & 0.85 & 11400/11300 \\
		2013-08-03 08:34 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.49 & 0.85 & 11400/11300 \\

		\hline\noalign{\smallskip}

	\end{tabular}

\end{table*}

@@

\begin{figure}
%%	\centering \includegraphics[width=9cm]{photometry_mosaic.pdf}
\caption{Mosaic showing all used imaging. The host of GRB~111117A is marked by
	a red circle with a 2\farc0 radius. This is the same size as the aperture used
	for deriving the photometry. Each panel is 20\arcsec~in size, north is up, east
	is left. Worth noting is the relative depth of the GMOS-N and FORS2 $z$-band
	images. For reference, the object located to the south-east of the host and
	marked with a blue, dashed circle has an extinction corrected magnitude of
	$23.11 \pm 0.09$ ($23.10 \pm 0.18$) in the FORS2 (GMOS-N) $z$-band.}
\label{fig:photplot}
\end{figure}

\begin{table*}[!ht]

	\centering
	\caption{Overview of the photometric observations. \label{tab:phot_overview}}
	\begin{tabular}{ccccccc}
		\hline\hline
		{Observation epoch} &  Exptime & Telescope/instrument & Filter & Airmass & Image quality & Host brightness\tablefootmark{a}  \\ [1.5pt]
		\hline
		({UT}) & ({ks}) &    & {} & & (arcsec)  & (AB mag)  \\ [1.5pt]
		\hline
		2013-08-30 07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
		2011-11-17 20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
		2011-11-17 20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
		2013-07-17 08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
		2011-11-28 21:10 & 3.60 & TNG/DOLORES & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
		2011-11-17 20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
		2013-08-03 09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
		2011-11-28 06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
		2013-07-13 09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
		2013-06-24 09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
		2013-06-27 09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
		2013-06-28 09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
		\hline\noalign{\smallskip}

	\end{tabular}

	\tablefoot{
		\tablefoottext{a}{All magnitudes are given in the AB system and are not corrected for the expected Galactic foreground extinction corresponding to a reddening of $E_{B-V}=0.027$\,mag.}}
\end{table*}

@@

This securely makes GRB~111117A, by far, the highest redshift sGRB detected to
date. The redshift and classification of GRB~111117A imply that it occurred when
the universe was younger by $\sim$ 3 Gyr compared to any other securely
classified sGRB ever detected. If the merger of NSs is the primary agent for the
$r$-process element enrichment of the universe \citep{Goriely2011, Ji2016,
	Komiya2016, Safarzadeh2017}, this marks the earliest detection of this process.

\begin{figure}
	\centering
%%	\includegraphics[width=9cm]{XLC_111117A_rest.pdf}
	\caption{Rest-frame XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shaded region is a compilation of long GRB lightcurves \citep{Evans2007, Evans2009} where the color represents density. The light blue lines are sGRB lightcurves from bursts with duration of $T_{90} \lesssim 2$ s and those that were classified as short in \citet{Kann2011, Berger2014, DAvanzo2014a}. The thick green line is GRB~111117A. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population. The insert shows the X-ray luminosity distributions of sGRBs and lGRBs at 10 ks, indicated by the vertical dashed line in the main panel.}
	\label{fig:sxray_lightcurve}
\end{figure}

@@

\begin{figure}
%%	\centering \includegraphics[width=9cm]{NH_z.pdf} 
\caption{Rest frame, X-ray derived equivalent hydrogen column density of
	GRB~111117A compared to complete samples of both long and short populations of
	GRBs. The sample of lGRBs from \citet{Arcodia2016} is shown in red, where
	detections are also shown with a kernel density estimate of the points and the
	limits on $N_\mathrm{H,X}$ are shown with arrows. The complete sample of sGRB by
	\citet{DAvanzo2014a} is shown in blue, where again the limits are indicated by
	arrows. Marginalizations over both axes are shown on the right and on the top of
	the plot, where the limits are shown as semi-transparent bars and detections as
	solid ones. The red curves in the marginalization plots are again the kernel
	density estimates of the \citet{Arcodia2016} sample.} \label{fig:NH_z}
\end{figure}

@@

A critical test to assess if the power law delay time models can be accommodated
by the current observation is to check if the implied sGRB rate at higher
redshift does not exceed the number of observed sGRBs without redshifts. Of the
100  sGRBs observed by \textit{Swift}, 20 have secure redshifts, and another 7
have a tentative redshift measurement\footnote{This is based on
	\url{http://www.astro.caltech.edu/grbox/grbox.php}, selecting all short,
	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning that $> 73$ per cent
of all sGRBs observed with \textit{Swift} are missing redshifts. More recently,
\citet{Fong2017} \citep{LIGOScientificCollaboration2017b} compiled a list of 36
(33) sGRBs with redshift measurements and, using this number, the redshift
incompleteness of sGRBs decreases to 64 per cent. Additional to the potential
number of high-$z$ event already detected but missing redshifts,
\citet{Behroozi2014} parametrized the \textit{Swift} redshift sensitivity and
found that the mean detection probability for sGRBs at $z \sim2$ was only
$\sim1$ per cent of the mean detection probability at $z \sim1$, assuming that
the unknown beaming angle of sGRBs stays constant with time. What this means is
that at the present, there is almost no limit on the number of sGRBs that could
be at redshifts $z > 2$.

@@

\begin{figure}
	\centering
%%	\includegraphics[width=9cm]{prog_sep.pdf}
	\caption{Constraints on the initial progenitor separation, given binary constituent masses. As can be seen, a heavier system will inspiral faster, leading to a weaker constraint on the initial separation of the binary. The most probable value, the 68 per cent, and 98 percent posterior predictive intervals of the NS binary mass distribution from \citet{Kiziltan2013} are shown with the dashed ellipses, which corresponds to a constraint on maximal initial separation of $a_0 <3.1^{+0.2(+0.4)}_{-0.2(-0.4)}~R_\odot$, respectively.}
	\label{fig:prog_sep}
\end{figure}

@@

The delay time between formation and explosion is well accommodated by the
models of \citet{Belczynski2006}, where a range of binary NS systems and their
decays are followed in different host potentials.  Some of the longest delay
times are, however excluded. Given late type nature of the host, only a small
subset of the longest population synthesis models violates the delay time
constraint derived here \citep{OShaughnessy2008}. The same holds, if NS binaries
are primarily formed through dynamical interactions in globular clusters
\citep{Lee2010, Church2011}.
